paper_replication

### 2021-12-16 paper_replication.Rmd
kable hold_position

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#load in relevant data
providers <- read.csv("provider_list_geo.csv")
countycentroids <- read.csv("county_pop_centroids.csv")
countydem <- read.csv("county_demographics.csv")
abortions <- read.csv("cleanedabortiondata.csv")
restrictions <- read.csv("state_restrictions2010.csv")
countypres <- read.csv("countypres.csv")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
########### clean county dem data ###########
countydem <- read.csv("county_demographics.csv")

@@

table_1 <- countydemanalysis %>% group_by(state) %>% summarize(totalcounties = n())
usedcounties <- aggregate(AbortionRate2010 ~ state, data = countydemanalysis, function(x) {sum(!(is.na(x)))}, na.action = NULL)
table_1 <- left_join(table_1, usedcounties, by = "state")
table_1 %>% mutate(percent = AbortionRate2010/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Abortion Data in 2010", "Percent"), caption = "Number of Counties with Abortion Data Included in Analyses by State", digits = 3) %>% kable_styling()

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Model Summaries (Note coefficients are per 1000 women)
table_3 <- tibble(Characteristic = "Highly Restrictive Legislation",
                  Beta = model_no_dist$coefficients[2]*1000,
                  "p-value" = c('< 0.001')) %>% 
  kbl(caption = "Change in Abortions per 1000 Women in Model without Distances", 
          digits = 3)
table_3
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
p <- summ_dist$coefficients[4,4] %>% round(3)
table_4 <- tibble(Characteristic = c("Highly Restrictive Legislation", "Distance to Provider",
                                     "Distance/Restriction Interaction"),
                  Beta = c(model_dist$coefficients[2]*1000, model_dist$coefficients[3]*1000,
                           model_dist$coefficients[4]*1000),
                  "p-value" = c('< 0.001', '< 0.001', p)) %>%
  kbl(caption = "Change in Abortions per 1000 Women in Model with Distances")
table_4

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#### Begin Extension #####
####### add in natality, maternalmortality, and infantdeath data #######
natality <- read.delim("natalitydata.txt")
maternalmortality <- read.delim("maternalmortality.txt")
infantdeath <- read.delim("infantdeathrecords.txt")

@@

```
## Expansion to Data from All States
```{r, echo=FALSE, message=FALSE, warning=FALSE}
#### recode restriction level and include all counties with abortion data in analysis ####
countyanalysis <- function(varname) {
  countydem <- countydem %>% mutate(hasdata = case_when(is.na(countydem[[varname]]) ~ 0, 
                                                              !(is.na(countydem[[varname]])) ~ 1))
  countydemanalysis <- countydem %>% filter(hasdata == 1)
  summarytable <- countydem %>% group_by(state) %>% summarize(totalcounties = n())
  usedcounties <- aggregate(countydem[[varname]] ~ state, data = countydem, function(x) {sum(!(is.na(x)))}, na.action = NULL)
  summarytable <- left_join(summarytable, usedcounties, by = "state")
  summarytable <- summarytable %>% mutate(percent = summarytable[, 3]/totalcounties) %>% 
    kbl(col.names = c("State", "Total Counties", "Counties with Data in 2010", "Percent"), 
        caption = "Number of Counties with Data Included in Analysis by State") %>% 
    kable_styling()

@@

# Model Summaries (Note coefficients are per 1000 women)
table_5 <- tibble(Characteristic = "Highly Restrictive Legislation",
                  Beta = all_state_model_no_dist$coefficients[2]*1000,
                  "p-value" = c('< 0.001')) %>% 
  kbl(caption = "All States Data: Change in Abortions per 1000 Women in Model without Distances", 
          digits = 3)
table_5

p <- all_state_model_dist$coefficients[4,4] %>% round(3)
table_6 <- tibble(Characteristic = c("Highly Restrictive Legislation", "Distance to Provider",
                                     "Distance/Restriction Interaction"),
                  Beta = c(all_state_model_dist$coefficients[2]*1000, 
                           all_state_model_dist$coefficients[3]*1000,
                           all_state_model_dist$coefficients[4]*1000),
                  "p-value" = c('< 0.001', '< 0.001', p)) %>%
  kbl(caption = "All States Data: Change in Abortions per 1000 Women in Model with Distances")
table_6

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
###further clean up maternaldeathrate, infantdeathrate, birthrate, and fertilityrate data ####
countydem$maternaldeathrate <- gsub(countydem$maternaldeathrate, pattern = "[()]", replacement = "")
countydem$maternaldeathrate <- gsub(countydem$maternaldeathrate, pattern = "Unreliable", replacement = "")
countydem$maternaldeathrate <- as.numeric(countydem$maternaldeathrate)
# hist(countydem$maternaldeathrate)

@@

## Extesion to Infant Death Rate, Birth Rate, and Fertility Rate

@@

# Model Summaries
table_7 <- tibble(Characteristic = "Highly Restrictive Legislation",
                  Beta = large_county_no_dist$coefficients[2],
                  "p-value" = c('< 0.001')) %>% 
  kbl(caption = "Large County Data: Change in Infant Death Rate per 1000 **** in Model without Distances", 
          digits = 4)
table_7

@@

table_8 <- tibble(Characteristic = c("Highly Restrictive Legislation", "Distance to Provider",
                                     "Distance/Restriction Interaction"),
                  Beta = c(large_county_dist$coefficients[2], 
                           large_county_dist$coefficients[3],
                           large_county_dist$coefficients[4]),
                  "p-value" = c('< 0.001', p_2, p_3)) %>%
  kbl(caption = "Large County Data: Change in Infant Death Rate per 1000 **** in Model with Distances")
table_8

@@

statelevelnatality <- read.delim("statelevelnatality.txt")
statelevelinfantdeaths <- read.delim("statelevelinfantdeaths.txt")
statelevelfetaldeaths <- read.delim("statelevelfetaldeaths.txt")
statelevelmaternalmortality <- read.delim("statelevelmaternalmortality.txt")

@@

p <- model_no_dist$coefficients[2,4] %>% round(3)
# Model Summaries
table_9 <- tibble(Characteristic = "Highly Restrictive Legislation",
                  Beta = model_no_dist$coefficients[2],
                  "p-value" = c(p)) %>% 
  kbl(caption = "State Data: Change in Birth Rate Rate per 1000 People in Model without Distances", 
          digits = 4)
table_9

@@

table_10 <- tibble(Characteristic = c("Highly Restrictive Legislation", "Distance to Provider",
                                     "Distance/Restriction Interaction"),
                  Beta = c(model_dist$coefficients[2], 
                           model_dist$coefficients[3],
                           model_dist$coefficients[4]),
                  "p-value" = c(p_1, p_2, p_3)) %>%
  kbl(caption = "State Data: Change in Birth Rate per 1000 People in Model with Distances")
table_10
```

@@

p <- model_no_dist$coefficients[2,4] %>% round(3)
# Model Summaries
table_no_dist <- tibble(Characteristic = "Highly Restrictive Legislation",
                  Beta = model_no_dist$coefficients[2],
                  "p-value" = c(p)) %>% 
  kbl(caption = "State Data: Change in Fertility Rate Rate per 1000 Women in Model without Distances", 
          digits = 4)
table_no_dist

@@

table_dist <- tibble(Characteristic = c("Highly Restrictive Legislation", "Distance to Provider",
                                     "Distance/Restriction Interaction"),
                  Beta = c(model_dist$coefficients[2], 
                           model_dist$coefficients[3],
                           model_dist$coefficients[4]),
                  "p-value" = c(p_1, p_2, p_3)) %>%
  kbl(caption = "State Data: Change in Fertility Rate per 1000 Women in Model with Distances")
table_dist
```

@@

# p <- model_no_dist$coefficients[2,4] %>% round(3)
# Model Summaries
table_no_dist <- tibble(Characteristic = "Highly Restrictive Legislation",
                  Beta = model_no_dist$coefficients[2],
                  "p-value" = '< 0.001') %>% 
  kbl(caption = "State Data: Change in Infant Death Rate Rate per 1000 **** in Model without Distances", 
          digits = 4)
table_no_dist

@@

table_dist <- tibble(Characteristic = c("Highly Restrictive Legislation", "Distance to Provider",
                                     "Distance/Restriction Interaction"),
                  Beta = c(model_dist$coefficients[2], 
                           model_dist$coefficients[3],
                           model_dist$coefficients[4]),
                  "p-value" = c('< 0.001', p_2, p_3)) %>%
  kbl(caption = "State Data: Change in Infant Death Rate per 1000 **** in Model with Distances")
table_dist
```

@@

p <- model_no_dist$coefficients[2,4] %>% round(3)
# Model Summaries
table_no_dist <- tibble(Characteristic = "Highly Restrictive Legislation",
                  Beta = model_no_dist$coefficients[2],
                  "p-value" = p) %>% 
  kbl(caption = "State Data: Change in Percent Fetal Deaths in Model without Distances", 
          digits = 4)
table_no_dist

@@

table_dist <- tibble(Characteristic = c("Highly Restrictive Legislation", "Distance to Provider",
                                     "Distance/Restriction Interaction"),
                  Beta = c(model_dist$coefficients[2], 
                           model_dist$coefficients[3],
                           model_dist$coefficients[4]),
                  "p-value" = c(p_1, p_2, p_3)) %>%
  kbl(caption = "State Data: Change in Percent Fetal Deaths in Model with Distances")
table_dist
```

@@

p <- model_no_dist$coefficients[2,4] %>% round(3)
# Model Summaries
table_no_dist <- tibble(Characteristic = "Highly Restrictive Legislation",
                  Beta = model_no_dist$coefficients[2],
                  "p-value" = p) %>% 
  kbl(caption = "State Data: Change in Maternal Deaths per ***** in Model without Distances", 
          digits = 4)
table_no_dist

@@

table_dist <- tibble(Characteristic = c("Highly Restrictive Legislation", "Distance to Provider",
                                     "Distance/Restriction Interaction"),
                  Beta = c(model_dist$coefficients[2], 
                           model_dist$coefficients[3],
                           model_dist$coefficients[4]),
                  "p-value" = c(p_1, p_2, p_3)) %>%
  kbl(caption = "State Data: Change in Percent Maternal Deaths per **** in Model with Distances")
table_dist
```

>>>>

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#load in relevant data
providers <- read.csv("data/provider_list_geo.csv")
countycentroids <- read.csv("data/county_pop_centroids.csv")
countydem <- read.csv("data/county_demographics.csv")
abortions <- read.csv("data/cleanedabortiondata.csv")
restrictions <- read.csv("data/state_restrictions2010.csv")
countypres <- read.csv("data/countypres.csv")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
########### clean county dem data ###########
countydem <- read.csv("data/county_demographics.csv")

@@

table_1 <- countydemanalysis %>% group_by(state) %>% summarize(totalcounties = n())
usedcounties <- aggregate(AbortionRate2010 ~ state, data = countydemanalysis, function(x) {sum(!(is.na(x)))}, na.action = NULL)
table_1 <- left_join(table_1, usedcounties, by = "state")
table_1 %>% mutate(percent = AbortionRate2010/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Abortion Data in 2010", "Percent"), caption = "Number of Counties with Abortion Data Included in Analyses by State", digits = 3) %>% kable_styling(latex_options = 'hold_position')

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Model Summaries (Note coefficients are per 1000 women)
table_3 <- tibble(Characteristic = "Highly Restrictive Legislation",
                  Beta = model_no_dist$coefficients[2]*1000,
                  "p-value" = c('< 0.001')) %>% 
  kbl(caption = "Change in Abortions per 1000 Women in Model without Distances", 
          digits = 3) %>%
  kable_styling(latex_options = 'hold_position')
table_3
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
p <- summ_dist$coefficients[4,4] %>% round(3)
table_4 <- tibble(Characteristic = c("Highly Restrictive Legislation", "Distance to Provider",
                                     "Distance/Restriction Interaction"),
                  Beta = c(model_dist$coefficients[2]*1000, model_dist$coefficients[3]*1000,
                           model_dist$coefficients[4]*1000),
                  "p-value" = c('< 0.001', '< 0.001', p)) %>%
  kbl(caption = "Change in Abortions per 1000 Women in Model with Distances") %>%
  kable_styling(latex_options = 'hold_position')
table_4

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#### Begin Extension #####
####### add in natality, maternalmortality, and infantdeath data #######
natality <- read.delim("data/natalitydata.txt")
maternalmortality <- read.delim("data/maternalmortality.txt")
infantdeath <- read.delim("data/infantdeathrecords.txt")

@@

```
## Expansion to Data from All States
```{r, echo=FALSE, message=FALSE, warning=FALSE}
#### recode restriction level and include all counties with abortion data in analysis ####
countyanalysis <- function(varname) {
  countydem <- countydem %>% mutate(hasdata = case_when(is.na(countydem[[varname]]) ~ 0, 
                                                              !(is.na(countydem[[varname]])) ~ 1))
  countydemanalysis <- countydem %>% filter(hasdata == 1)
  summarytable <- countydem %>% group_by(state) %>% summarize(totalcounties = n())
  usedcounties <- aggregate(countydem[[varname]] ~ state, data = countydem, function(x) {sum(!(is.na(x)))}, na.action = NULL)
  summarytable <- left_join(summarytable, usedcounties, by = "state")
  summarytable <- summarytable %>% mutate(percent = summarytable[, 3]/totalcounties) %>% 
    kbl(col.names = c("State", "Total Counties", "Counties with Data in 2010", "Percent"), 
        caption = "Number of Counties with Data Included in Analysis by State") %>% 
    kable_styling(latex_options = 'hold_position')

@@

# Model Summaries (Note coefficients are per 1000 women)
table_5 <- tibble(Characteristic = "Highly Restrictive Legislation",
                  Beta = all_state_model_no_dist$coefficients[2]*1000,
                  "p-value" = c('< 0.001')) %>% 
  kbl(caption = "All States Data: Change in Abortions per 1000 Women in Model without Distances", 
          digits = 3) %>%
  kable_styling(latex_options = 'hold_position')
table_5

p <- all_state_model_dist$coefficients[4,4] %>% round(3)
table_6 <- tibble(Characteristic = c("Highly Restrictive Legislation", "Distance to Provider",
                                     "Distance/Restriction Interaction"),
                  Beta = c(all_state_model_dist$coefficients[2]*1000, 
                           all_state_model_dist$coefficients[3]*1000,
                           all_state_model_dist$coefficients[4]*1000),
                  "p-value" = c('< 0.001', '< 0.001', p)) %>%
  kbl(caption = "All States Data: Change in Abortions per 1000 Women in Model with Distances") %>%
  kable_styling(latex_options = 'hold_position')
table_6

@@

## Extension to Infant Death Rate, Birth Rate, and Fertility Rate
```{r, echo=FALSE, message=FALSE, warning=FALSE}
###further clean up maternaldeathrate, infantdeathrate, birthrate, and fertilityrate data ####
countydem$maternaldeathrate <- gsub(countydem$maternaldeathrate, pattern = "[()]", replacement = "")
countydem$maternaldeathrate <- gsub(countydem$maternaldeathrate, pattern = "Unreliable", replacement = "")
countydem$maternaldeathrate <- as.numeric(countydem$maternaldeathrate)
# hist(countydem$maternaldeathrate)

@@

@@

# Model Summaries
table_7 <- tibble(Characteristic = "Highly Restrictive Legislation",
                  Beta = large_county_no_dist$coefficients[2],
                  "p-value" = c('< 0.001')) %>% 
  kbl(caption = "Large County Data: Change in Infant Death Rate per 1000 **** in Model without Distances", 
          digits = 4) %>%
  kable_styling(latex_options = 'hold_position')
table_7

@@

table_8 <- tibble(Characteristic = c("Highly Restrictive Legislation", "Distance to Provider",
                                     "Distance/Restriction Interaction"),
                  Beta = c(large_county_dist$coefficients[2], 
                           large_county_dist$coefficients[3],
                           large_county_dist$coefficients[4]),
                  "p-value" = c('< 0.001', p_2, p_3)) %>%
  kbl(caption = "Large County Data: Change in Infant Death Rate per 1000 **** in Model with Distances") %>%
  kable_styling(latex_options = 'hold_position')
table_8

@@

statelevelnatality <- read.delim("data/statelevelnatality.txt")
statelevelinfantdeaths <- read.delim("data/statelevelinfantdeaths.txt")
statelevelfetaldeaths <- read.delim("data/statelevelfetaldeaths.txt")
statelevelmaternalmortality <- read.delim("data/statelevelmaternalmortality.txt")

@@

p <- model_no_dist$coefficients[2,4] %>% round(3)
# Model Summaries
table_9 <- tibble(Characteristic = "Highly Restrictive Legislation",
                  Beta = model_no_dist$coefficients[2],
                  "p-value" = c(p)) %>% 
  kbl(caption = "State Data: Change in Birth Rate Rate per 1000 People in Model without Distances", 
          digits = 4) %>%
  kable_styling(latex_options = 'hold_position')
table_9

@@

table_10 <- tibble(Characteristic = c("Highly Restrictive Legislation", "Distance to Provider",
                                     "Distance/Restriction Interaction"),
                  Beta = c(model_dist$coefficients[2], 
                           model_dist$coefficients[3],
                           model_dist$coefficients[4]),
                  "p-value" = c(p_1, p_2, p_3)) %>%
  kbl(caption = "State Data: Change in Birth Rate per 1000 People in Model with Distances") %>%
  kable_styling(latex_options = 'hold_position')
table_10
```

@@

p <- model_no_dist$coefficients[2,4] %>% round(3)
# Model Summaries
table_no_dist <- tibble(Characteristic = "Highly Restrictive Legislation",
                  Beta = model_no_dist$coefficients[2],
                  "p-value" = c(p)) %>% 
  kbl(caption = "State Data: Change in Fertility Rate Rate per 1000 Women in Model without Distances", 
          digits = 4) %>%
  kable_styling(latex_options = 'hold_position')
table_no_dist

@@

table_dist <- tibble(Characteristic = c("Highly Restrictive Legislation", "Distance to Provider",
                                     "Distance/Restriction Interaction"),
                  Beta = c(model_dist$coefficients[2], 
                           model_dist$coefficients[3],
                           model_dist$coefficients[4]),
                  "p-value" = c(p_1, p_2, p_3)) %>%
  kbl(caption = "State Data: Change in Fertility Rate per 1000 Women in Model with Distances") %>%
  kable_styling(latex_options = 'hold_position')
table_dist
```

@@

# p <- model_no_dist$coefficients[2,4] %>% round(3)
# Model Summaries
table_no_dist <- tibble(Characteristic = "Highly Restrictive Legislation",
                  Beta = model_no_dist$coefficients[2],
                  "p-value" = '< 0.001') %>% 
  kbl(caption = "State Data: Change in Infant Death Rate Rate per 1000 **** in Model without Distances", 
          digits = 4) %>%
  kable_styling(latex_options = 'hold_position')
table_no_dist

@@

table_dist <- tibble(Characteristic = c("Highly Restrictive Legislation", "Distance to Provider",
                                     "Distance/Restriction Interaction"),
                  Beta = c(model_dist$coefficients[2], 
                           model_dist$coefficients[3],
                           model_dist$coefficients[4]),
                  "p-value" = c('< 0.001', p_2, p_3)) %>%
  kbl(caption = "State Data: Change in Infant Death Rate per 1000 **** in Model with Distances") %>%
  kable_styling(latex_options = 'hold_position')
table_dist
```

@@

p <- model_no_dist$coefficients[2,4] %>% round(3)
# Model Summaries
table_no_dist <- tibble(Characteristic = "Highly Restrictive Legislation",
                  Beta = model_no_dist$coefficients[2],
                  "p-value" = p) %>% 
  kbl(caption = "State Data: Change in Percent Fetal Deaths in Model without Distances", 
          digits = 4) %>%
  kable_styling(latex_options = 'hold_position')
table_no_dist

@@

table_dist <- tibble(Characteristic = c("Highly Restrictive Legislation", "Distance to Provider",
                                     "Distance/Restriction Interaction"),
                  Beta = c(model_dist$coefficients[2], 
                           model_dist$coefficients[3],
                           model_dist$coefficients[4]),
                  "p-value" = c(p_1, p_2, p_3)) %>%
  kbl(caption = "State Data: Change in Percent Fetal Deaths in Model with Distances") %>%
  kable_styling(latex_options = 'hold_position')
table_dist
```

@@

p <- model_no_dist$coefficients[2,4] %>% round(3)
# Model Summaries
table_no_dist <- tibble(Characteristic = "Highly Restrictive Legislation",
                  Beta = model_no_dist$coefficients[2],
                  "p-value" = p) %>% 
  kbl(caption = "State Data: Change in Maternal Deaths per ***** in Model without Distances", 
          digits = 4) %>%
  kable_styling(latex_options = 'hold_position')
table_no_dist

@@

table_dist <- tibble(Characteristic = c("Highly Restrictive Legislation", "Distance to Provider",
                                     "Distance/Restriction Interaction"),
                  Beta = c(model_dist$coefficients[2], 
                           model_dist$coefficients[3],
                           model_dist$coefficients[4]),
                  "p-value" = c(p_1, p_2, p_3)) %>%
  kbl(caption = "State Data: Change in Percent Maternal Deaths per **** in Model with Distances") %>%
  kable_styling(latex_options = 'hold_position')
table_dist
```
